---
title: "Analysis of eBay Auction Data"
author: "Esham Bin Rashid"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
  pdf_document:
    toc: true
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    fig_caption: true
---

```{r setup, include=FALSE}
library(knitr)
knitr::knit_hooks$set(append_name = function(before, options) {
  if (before) {
    return("")
  } else {
    return(paste("<div style='text-align: left; font-style: italic;'>- Esham Bin Rashid</div>", sep="\n"))
  }
})
opts_chunk$set(append_name = TRUE,echo = TRUE, message = FALSE, warning = FALSE)
```

# Abstract

The objective is to build a model to classify eBay auctions as competitive or non-competitive based on a dataset detailing auctions from May–June 2004.

# Introduction

This analysis involves pre-processing the eBay auction dataset, applying classification models, and discussing their performances to predict auction competitiveness.

# Data Loading and Preprocessing

```{r}
library(tidyverse)
library(randomForest)
library(caret)
library(pROC)
library(ggplot2)
```

```{r}
# Loading the data
eBayAuctions_EBR <- read.csv("eBayAuctions.csv")
# Converting 'Duration' to a categorical variable
eBayAuctions_EBR$Duration <- as.factor(eBayAuctions_EBR$Duration)
# Converting 'Competitive.' to a cateforical variables with two levels
eBayAuctions_EBR$Competitive. <- factor(eBayAuctions_EBR$Competitive., levels = c(0,1), labels = c("Noncompetitive", "Competitive"))
# Checking the structure of the dataframe
str(eBayAuctions_EBR)
```

# Auction Data Summary

```{r}
summary(eBayAuctions_EBR)
```

# Methodology

For this analysis random forest classifier for this task due to its robustness and ability to handle both numerical and categorical data effectively.

# Model Building and Validation

```{r}
set.seed(2024)

# Splitting the data into training and validation sets
training_index <- createDataPartition(eBayAuctions_EBR$Competitive., p = 0.75, list = FALSE)
training_data <- eBayAuctions_EBR[training_index, ]
validation_data <- eBayAuctions_EBR[-training_index, ]
```

```{r}
# Training the random forest model
rf_model <- randomForest(Competitive. ~ ., data = training_data, ntree = 100)
# Viewing the random forest model
rf_model
```

# Model Evaluation

```{r}
# Prediction and confusion matrix on training data
training_preds <- predict(rf_model, training_data)
confusionMatrix(training_preds, training_data$Competitive.)
```

```{r}
# Prediction and confusion matrix on validation data
validation_preds <- predict(rf_model, validation_data)
confusionMatrix(validation_preds, validation_data$Competitive.)
```

# Graphical Evaluation

```{r}
prob <- predict(rf_model, validation_data, type = "prob")[, 2]
lift_df <- data.frame(actual = validation_data$Competitive, prob = prob)
lift_obj <- lift(lift_df$actual ~ lift_df$prob)

plot(lift_obj, main = "Lift Chart", xlab = "Percentile", ylab = "Lift")
```

# Conclusion

After analyzing the confusion matrix and the lift chart I can confidently say that the model has a good predictive power, especially in the top percentiles that means, the model is quite effective in ranking the cases by their likelihood of being a positive instance that is being competitive in the auction.By using this model ebay can predict whether an auction is going to competitive or not based on the given variables.

Report Made By - Esham Bin Rashid, Student ID: 00864931, Graduate Student, MS in Business Analytics, University of New Haven
